Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a High Performance Database System \\

### 2018-07-01 Presentations/Paper/paper_thread_pool.tex

\title{Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a high performance Database System \\
}

@@

\subsection{Comparison against Baseline Implementation}
As most important aspect of our evaluation, we compare our thread pool against an approximated version of the baseline implementation. Since BOLSTER is deeply integrated into geckoDB and the workload and usage of multithreading depends on the graph data in the system, we can not directly use the BOLSTER implementation for our evaluation. Thus, we extracted the respecting code parts into an own implementation and use this for a direct comparison against the thread pool. In Figure \ref{fig4}, we show the approximated version of the BOLSTER multithreading code that we use as baseline implementation. In our benchmark. the method \emph{cmp\_baseline} is called multiple times with a different \emph{numThreads} parameter. Within each call, \emph{numThreads} threads are created and all of these execute the benchmark method v, that calculates a huge amount of exponential values in a loop. After the calling thread takes part in execution, it waits until all created threads are joined. The elapsed processing time is measured from the begin of the method until joining all threads finished.

@@

In our performance comparison, both implementations are executed three times with $4, 8, 16, 32$ as \emph{numThreads} parameter and the summed results are divided by three. We consider a number between $4$ and $32$ as a useful amount of threads to reach a sufficient performance increase due to parallelization, but also to have the possibility to debug and profile them.

In Figure \ref{fig6}, we show the comparison results of both mentioned implementations and the different \emph{numThreads} parameters.

@@

>>>>

\title{Thread Pool in GeckoDB/BOLSTER - Implementing a Thread Pool for a High Performance Database System \\}

@@

\subsection{TODO: Validity of Results}

\subsection{Comparison against Baseline Implementation}
As most important aspect of our evaluation, we compare our thread pool against an approximated version of the baseline implementation. To justify the performance comparison results, we present the different benchmark implementations in detail before we show the evaluation results.

\subsubsection{Benchmark Implementations}
Since BOLSTER is deeply integrated into geckoDB and the workload and usage of multithreading depends on the graph data in the system, we can not directly use the BOLSTER implementation for our evaluation. Thus, we extracted the respecting code parts into an own implementation and use this for a direct comparison against the thread pool.

In Figure \ref{fig4}, we show the approximated version of the BOLSTER multithreading code that we use as baseline implementation. In our benchmark. the method \emph{cmp\_baseline} is called multiple times with a different \emph{numThreads} parameter. Within each call, \emph{numThreads} threads are created and all of these execute the benchmark method v, that calculates a huge amount of exponential values in a loop. After the calling thread takes part in execution, it waits until all created threads are joined. The elapsed processing time is measured from the begin of the method until joining all threads finished.

@@

\subsubsection{Comparison Results}
In our performance comparison, both implementations are executed three times with $4, 8, 16, 32$ as \emph{numThreads} parameter and the summed results are divided by three. We consider a number between $4$ and $32$ as a useful amount of threads to reach a sufficient performance increase due to parallelization, but also to have the possibility to debug and profile them.

In Figure \ref{fig6}, we show the comparison results of both mentioned implementations and the different \emph{numThreads} parameters. Our thread pool implementation is faster than the baseline for all \emph{numThreads} parameters. Based on Figure \ref{fig6}, we make the following conclusions.

Firstly, compared to strict processing workload, create, and later on join threads, consumes much time. Our thread pool implementation reaches up to a multiple of performance increase compared to the baseline. Since the creation and join of threads is the main difference between both tests, reusing threads will lead to this noticeable performance increase.

Secondly, the task creation and enqueuing does not have heavy impact on the performance, since the difference in execution time \textbf{TODO: Complete!}

@@

\subsubsection{Summary}
As mentioned before, our implementation reaches better performance than the baseline regarding our specific tests.

\subsection{Waiting Time Evaluation}
\begin{figure}
	\includegraphics[width=0.5\textwidth]{img/pool_avg.png}
	\caption{Time measures in nanoseconds of thread pool (blue) and baseline (orange) implementation with different numbers of threads.  }
	\label{fig7}
\end{figure}
